---
title: "Group 1 Project"
subtitle: "Rayshader Package"
author: "Diego, Manroop, and Jessica"
date: "(`r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      beforeInit: "macros.js"
      ratio: '15:10'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
<div style = "position:fixed; visibility: hidden">
$$\require{color}\definecolor{yellow}{rgb}{1, 0.8, 0.16078431372549}$$
$$\require{color}\definecolor{orange}{rgb}{0.96078431372549, 0.525490196078431, 0.203921568627451}$$
$$\require{color}\definecolor{green}{rgb}{0, 0.474509803921569, 0.396078431372549}$$
</div>

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      yellow: ["{\\color{yellow}{#1}}", 1],
      orange: ["{\\color{orange}{#1}}", 1],
      green: ["{\\color{green}{#1}}", 1]
    },
    loader: {load: ['[tex]/color']},
    tex: {packages: {'[+]': ['color']}}
  }
});
</script>

<style>
.yellow {color: #FFCC29;}
.orange {color: #F58634;}
.green {color: #007965;}
</style>


```{r flair_color, echo=FALSE}
library(flair)
yellow <- "#FFCC29"
orange <- "#F58634"
green <- "#007965"
```


---

```{r, echo=FALSE, message=FALSE}

library(magick)
imagerayshade <- image_read("https://github.com/allisonhorst/stats-illustrations/blob/master/rstats-artwork/rayshader.png?raw=true")

imagerayshade
```

---

## Outline

---

## What is Rayshader?

---
## Introduction

Rayshader is a package used for 3D and 2D visualization. It’s functions can be used to create 3D and 2D maps and ggplots!  It does so through a series of overlays that when applied on top of each other form a 2D/3D map or plot. Rayrender (a subfeature of rayshader) can also be used to rotate the maps and show your creations from all angles as well as render your creation. 
---


## Install Packages

```{r}
# install.packages("rayshader")
# install.packages("crosstalk")
# install.packages("manipulateWidget")
```

---
## Load Libraries
```{r, message=FALSE, warning=FALSE}
library(rayshader)
library(here) #for unbreakable file paths
library(tidyverse)
library(raster) #handles raster files
```

---

## Where to get Topographic data
.pull-left[
[SRTM](https://dwtkns.com/srtm30m/)
```{r}
 sat<- image_read(here("Data", "srtm.png"))
sat
```

]

.pull-right[
[OpenTopo](https://portal.opentopography.org/datasets)
```{r}
topo<-image_read(here("Data", "opentopo.png"))
topo
```

]
---

## Load Data

---
## Examples:

.pull-left[
```{r quakes_2d, message=FALSE, warning=FALSE}

data("quakes")

quakes_sample <- quakes[seq(1,1000,by = 100),]

quakes_plot<-ggplot(data = quakes_sample, aes(x = lat, y = long)) + 
  geom_point(aes(size = mag, colour = mag)) +
  guides(size = F) + 
  labs(x = "Latitude", y = "Longitude") + 
  scale_colour_viridis_b(option = "B") + 
  scale_size_continuous(range = c(1,10))
```
]
.pull-right[
```{r}
quakes_plot
```
]
---
